---
Title: HW 5
Author: Yetsy Santa
Date: Feb 29, 2024
---

library(ggplot2)
library(tidyverse)
require(class)
require(caret)
load("BRFSS2022_rev.RData")
brfss22$Age_midpt <- fct_recode(brfss22$X_AGEG5YR, "21" = "Age 18 to 24",
                                "27" = "Age 25 to 29", "32" = "Age 30 to 34",
                                "37" = "Age 35 to 39", "42" = "Age 40 to 44",
                                "47" = "Age 45 to 49", "52" = "Age 50 to 54",
                                "57" = "Age 55 to 59", "62" = "Age 60 to 64",
                                "67" = "Age 65 to 69", "72" = "Age 70 to 74",
                                "77" = "Age 75 to 79", "82" = "Age 80 or older",
                                NULL = "Dont know/Refused/Missing")
brfss22$Age_midpt <- as.numeric(levels(brfss22$Age_midpt))[brfss22$Age_midpt]
brfss22$Educ_number <- fct_recode(brfss22$EDUCA, 
                                  "0" = "Never attended school or only kindergarten", 
                                  "4.5" = "Grades 1 through 8 (Elementary)",
                                  "10" = "Grades 9 through 11 (Some high school)",
                                  "12" = "Grade 12 or GED (High school graduate)",
                                  "14" = "College 1 year to 3 years (Some college or technical school)",
                                  "16" = "College 4 years or more (College graduate)",
                                  NULL = "Refused" )
brfss22$Educ_number <- as.numeric(levels(brfss22$Educ_number))[brfss22$Educ_number]

ACEdidntask <- (as.numeric(is.na(brfss22$ACEDEPRS)) + 
                  as.numeric(is.na(brfss22$ACEDRINK)) +
                  as.numeric(is.na(brfss22$ACEDRUGS)) +
                  as.numeric(is.na(brfss22$ACEPRISN)) +
                  as.numeric(is.na(brfss22$ACEDIVRC)) +
                  as.numeric(is.na(brfss22$ACEPUNCH)) +
                  as.numeric(is.na(brfss22$ACEHURT1)) +
                  as.numeric(is.na(brfss22$ACESWEAR)) +
                  as.numeric(is.na(brfss22$ACETOUCH)) )
select_ACE <- (ACEdidntask == 0) & !is.na(brfss22$MENTHLTH) # with zero missing values for any of ACE questions and not missing MENTLHLTH
brfss_ACE <- subset(brfss22, select_ACE)



summary(brfss_ACE$MENTHLTH[brfss_ACE$ACETOUCH == "Adverse Childhood Exper, never:  How often did anyone at least 5 years older than you or an adult, ever touch you sexually"]) 


summary(brfss_ACE$MENTHLTH[brfss_ACE$ACETOUCH == "once"])

summary(brfss_ACE$MENTHLTH[brfss_ACE$ACETOUCH == "more than once"])

sd(brfss_ACE$MENTHLTH[brfss_ACE$ACETOUCH == "Adverse Childhood Exper, never:  How often did anyone at least 5 years older than you or an adult, ever touch you sexually"], na.rm = TRUE)

sd(brfss_ACE$MENTHLTH[brfss_ACE$ACETOUCH == "once"], na.rm = TRUE)

sd(brfss_ACE$MENTHLTH[brfss_ACE$ACETOUCH == "more than once"], na.rm = TRUE)

summary(brfss_ACE$ACETOUCH) # N in each group




# is there an easier way?!?
library(plyr)
summary1 <- ddply(brfss_ACE,.(ACETOUCH), summarize, mean_mentalhealth = mean(MENTHLTH), 
                  sd_mentalhealth = sd(MENTHLTH), n_obs = sum(!is.na(MENTHLTH)) )
summary1






summary2 <- ddply(brfss_ACE,.(MENTHLTH >0), summarize, 
                  zero_ACETOUCH = sum(ACETOUCH == "Adverse Childhood Exper, never:  How often did anyone at least 5 years older than you or an adult, ever touch you sexually"), 
                  once_ACETOUCH = sum(ACETOUCH == "once"), 
                  mult_ACETOUCH = sum(ACETOUCH == "more than once") )
summary2


```{r}
brfss_ACE$ACETOUCH_recode <- fct_recode(brfss_ACE$ACETOUCH, 
                                        "0" = "Adverse Childhood Exper, never:  How often did anyone at least 5 years older than you or an adult, ever touch you sexually",
                                "0.5" = "once", 
                                "1" = "more than once",
                                NULL = "dont know not sure",
                                NULL = "refused"
)
brfss_ACE$ACEHURT_recode <- fct_recode(brfss_ACE$ACEHURT1, 
                                        "0" = "Adverse Childhood Exper, never: Not including spanking, (before age 18), how often did a parent or adult in your home ever hit, beat, kick, or physically hurt you in any way",
                                "0.5" = "once", 
                                "1" = "more than once",
                                NULL = "dont know not sure",
                                NULL = "refused"
)
brfss_ACE$ACETOUCH_recode <- as.numeric(levels(brfss_ACE$ACETOUCH_recode))[brfss_ACE$ACETOUCH_recode]
brfss_ACE$ACEHURT_recode <- as.numeric(levels(brfss_ACE$ACEHURT_recode))[brfss_ACE$ACEHURT_recode]

brfss_ACE$MENTHLTH_recode <- cut(brfss_ACE$MENTHLTH, breaks = c(-1,0,1,5,10,15,31))
summary(brfss_ACE$MENTHLTH_recode)
```




```{r}
# create a function to standardize
standardize_varb_to01 <- function(X_in) {
  (X_in - min(X_in, na.rm = TRUE))/( max(X_in, na.rm = TRUE) - min(X_in, na.rm = TRUE)  )
}
```



```{r}
X1 <- standardize_varb_to01(brfss_ACE$Age_midpt)
X2 <- standardize_varb_to01(brfss_ACE$Educ_number)
X3 <- brfss_ACE$ACETOUCH_recode
X4 <- brfss_ACE$ACEHURT_recode
# you could add more X variables...
Y <- brfss_ACE$MENTHLTH_recode

nonmissingobs <- complete.cases(Y,X1,X2,X3,X4)

X1 <- subset(X1, nonmissingobs)
X2 <- subset(X2, nonmissingobs)
X3 <- subset(X3, nonmissingobs)
X4 <- subset(X4, nonmissingobs)
dat_use <- data.frame(X1,X2,X3,X4)
Y <- subset(Y, nonmissingobs)
```


```{r}
set.seed(1234)
NN_obs <- length(Y)
select1 <- (runif(NN_obs) < 0.6)
train_data <- subset(dat_use,select1)
test_data <- subset(dat_use,(!select1))
cl_data <- Y[select1]
true_data <- Y[!select1]
```


```{r}
for (indx in seq(1, 9, by= 2)) {
 pred_y <- knn3Train(train_data, test_data, cl_data, k = indx, l = 0, prob = FALSE, use.all = TRUE)
 num_correct_labels <- sum(pred_y == true_data)
 correct_rate <- num_correct_labels/length(true_data)
 print(c(indx,correct_rate))
}
```








